#Reproducible Research: Peer Assignment 1
-----------------------------------------
#Using echo = TRUE ## someone else will be able to read the code.
#Loading and preprocessing the data
#1.Load the data
```{r}
unzip(zipfile="activity.zip")
activity <- "./activity.csv"
```{r}

#2.Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
data <- read.csv(activity, colClasses = c("integer", "Date", "factor"))
data$month <- as.numeric(format(data$date, "%m"))
str(data)
```{r}
    
#'data.frame':  17568 obs. of  3 variables:
```{r}
head(data)
```{r}

#1.What is mean total number of steps taken per day?
#For this part of the assignment, you can ignore the missing values in the dataset
```{r}
stepsnumber <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
```{r}
head(stepsnumber)
```{r}
         
#2.If you do not understand the difference between a histogram and a barplot, research #the difference between them. Make a histogram of the total number of steps taken each #day
```{r}
library(ggplot2)
qplot(stepsnumber, binwidth=800, xlab="total number of steps taken each day")
```{r}

#3.Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean(stepsnumber, na.rm=TRUE)
```{r}

```{r}
median(stepsnumber, na.rm=TRUE)
```{r}

#What is the average daily activity pattern?
#1.Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the #average number of steps taken, averaged across all days (y-axis)
```{r}
average <- aggregate(x=list(steps=data$steps), by=list(interval=as.numeric(as.character(data$interval))),FUN = mean, na.rm = TRUE)
ggplot(data = average, aes(x = interval, y = steps)) + geom_line(color = "steelblue", size = 1) +labs(title = "Time Series Plot of the 5-minute Interval")+ xlab("5-minute interval") + ylab("average number of steps taken")
```{r}

#2.Which 5-minute interval, on average across all the days in the dataset, contains #the maximum number of steps?
```{r}
average[which.max(average$steps), ]
```{r}

#Imputing missing values
#Note that there are a number of days/intervals where there are missing values (coded #as NA). The presence of missing days may introduce bias into some calculations or #summaries of the data.
#1.Calculate and report the total number of missing values in the dataset (i.e. the #total number of rows with NAs)
```{r}
missingvalues <- is.na(data$steps)
```{r}
table(missingvalues)
```{r}

#There are 2304 missing values.
#Sum(15264+2304) = 17568 obs.

#2.Devise a strategy for filling in all of the missing values in the dataset. The #strategy does not need to be sophisticated. For example, you could use the mean
#/median for that day, or the mean for that 5-minute interval, etc.
#To solve this problem, I will use the mean for that 5-minute interval to fill each NA #value in the steps column.
```{r}
fill <- function(steps, interval) {
    filled <- NA
   if (!is.na(steps)) 
      filled <- c(steps) else filled <- (average[average$interval == interval, "steps"])
   return(filled)
}
```{r}
filledata <- data
filledatafinal <- mapply(fill, filledata$steps, filledata$interval)
```{r}

#3.Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
newdata <- data 
for (i in 1:nrow(newdata)) {
if (is.na(newdata$steps[i])) {
newdata$steps[i] <- average[which(newdata$interval[i] == average$interval), ]$steps
       }
} 
```{r}
head(newdata)
```{r}
 
```{r}
sum(is.na(newdata))
```{r}

#4.Make a histogram of the total number of steps taken each day and Calculate and #report the mean and median total number of steps taken per day. Do these values #differ from the estimates from the first part of the assignment? What is the impact #of imputing missing data on the estimates of the total daily number of steps?
#Are there differences in activity patterns between weekdays and weekends?
```{r}
fillednewdata <- tapply(filledatafinal, filledata$date, FUN=sum)
qplot(fillednewdata, binwidth=1000, xlab="total number of steps taken each day")
```{r}
 mean(fillednewdata)
```{r}

```{r}
 median(fillednewdata)
```{r}


#We can observe that these median and mean values are higher than the other ones from #the initial dataset. This can be explained as in the original data there are some #days in which "steps" have missing data values (that means equal to 0). After filling #these missing values with "steps mean of associated interval value","0 values" #disappear and therefore, the median and mean values, increase.  

#For this part the weekdays() function may be of some help here. Use the dataset with #the filled-in missing values for this part.
#1.Create a new factor variable in the dataset with two levels – “weekday” and #“weekend” indicating whether a given date is a weekday or weekend day.
```{r}
newdata$dateType<-ifelse(as.POSIXlt(newdata$date)$wday %in% c(0,6), 'weekend', 'weekday')
```{r}
 averagenewdata <-aggregate(steps ~ interval + dateType, data=newdata, mean)
head(averagenewdata)
```{r}

#2.Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute #interval (x-axis) and the average number of steps taken, averaged across all weekday #days or weekend days (y-axis). See the README file in the GitHub repository to see an #example of what this plot should look like using simulated data.
```{r}
str(averagenewdata)
```{r}

```{r}
averagenewdata$interval <- as.numeric(as.character(averagenewdata$interval))
```{r}
ggplot(data=averagenewdata, aes(x=interval, y=steps,ymax = max(steps)*1.05)) +geom_line(aes(interval, steps))+facet_grid(dateType ~ .)+xlab("5-minute interval") + ylab("Number of steps")
```{r}
